---
title: "Analyze Effect of Mandatory Jail on Alcohol-involved Vehicle Fatality Rate"
output:
  pdf_document: 
    df_print: paged
    number_sections: yes
---

<style type="text/css">


Headers{/* Normal  */ font-size: 20pt;
}

body{ /* Normal  */
      font-size: 18px;
  }

</style>

```{r , include=FALSE}
knitr::opts_chunk$set(echo=FALSE,message=FALSE,warning=FALSE)
options(width = 120)
```


```{r, include=FALSE}
library(AER)
library(knitr)
library(ggplot2)
library(foreign)
library(table1)
library(dplyr)
library(kableExtra)
library(gridExtra)
library(MatchIt)
library(dagitty)
library(ggdag)
data('Fatalities')
```

# Introduction

This project mainly investigates the impact of mandatory jail on alcohol-involved vehicle fatality rate using the fix-effect model with data from the 48 states over 1982 through 1988, acquired from dataset Fatalities of AER package in R. This is the U.S. traffic fatalities panel data containing 336 observations on 34 variables. In the descriptive analysis, we noticed that mandatory jail seems to have a higher fatality rate. Then, we built the fix-effect model using alcohol-involved vehicle fatalities rate (afatal rate) against mandatory jail with year effect, state effect, and other 13 variables we controlled. However, after model diagnostics, it seemed that the assumptions we made in this model were violated. Finally, we considered using a propensity score to help explain the causal effect of mandatory jail. It was indicated that there was a causal effect for mandating jail to the fatality rate, and the mandating jail caused a higher fatality rate.  

Our main question of interest in this project is whether or not there is a significant effect of mandatory jail on alcohol-involved fatalities rate, and if so, how does it affect the rate.  

# Descriptive Analysis 
## Data Processing 
* \textit{\textbf{Missing  Value}}  After the first review of the data, we found there were two missing values, 'jail' and 'service' for California in 1988. From Alcohol Policies and Highway Vehicle Fatalities (Ruhm,1995), we knew that these two variables referred to legislation mandating jail or community service for the first Driving Under the Influence (DUI) conviction. By checking Digest of State Alcohol-Highway Safety Related Legislation (current as of Jan 1st 1988 and current as of Jan 1st 1989), in which stated that there were nither mandatory jail nor community service, we could assume that both of the missing value was 'no' and updated the dataset. With the missing data added, we could deal with a balanced panel.
* \textit{\textbf{Transformation of drinkage }}  We also noticed that some of the records of the drinkage were not an integer. Since the drinkage is an integer in alcohol laws of the United States, as far as we considered, it'll better we kept the same form. The following transformation method was applied: if new laws on minimum drink age were enforced in one year, the 'drinking' is the minimum drinkage as of the beginning of that year according to 'U.S. history of alcohol minimum purchase age by the state.' This adaptation allowed us to deal with fewer levels of factors for the drinkage.
* \textit{\textbf{Transformation of Response varible }}  Since most of the variables were related to alcohol policies, it is reasonable to think about their effect on the alcohol-involved fatalities. Thus, the response and measure of traffic deaths we used in this project is the alcohol-involved vehicle fatalities rate(afatal rate), which is the number of annual alcohol-involved traffic deaths per 10,000 people in the population in that state. 
* \textit{\textbf{Choose variables }}  Apart from mandatory jail, when choosing other variables, we first referred to Alcohol Policies and Highway Vehicle Fatalities(Ruhm,1995), in which a detailed summary of the results of previous relevant research was listed. From these results, we concluded that the following variables were related to fatality rate: breath test, beer tax, drink age, spirit consumption. Since we were given a more diversified data set, we also considered other variables. From the meaning of the variable, commonsense and pairwise scatterplots, we though the following variables might also be relevant: mandatory service, unemployment rate, income, percent of southern baptist, percent of Mormon, percent residing in "dry" counties, the proportion of drivers between 15 and 24, total vehicle miles, and GSP. Thus, in this project, in addition to the year and state effects, we also controlled for the above 13 variables.
```{r,results = "asis", message=FALSE, echo=FALSE, include=FALSE}
#initial data
F1<-Fatalities[, !names(Fatalities) %in%c('pop2124','pop1820','pop1517','fatal','nfatal2124','fatal2124','nfatal1820','fatal1820','nfatal1517','fatal1517','sfatal','nfatal','unempus','emppopus')]
F1[is.na(F1)]<-'no'
F1<-cbind(F1[, !names(F1)%in%c('afatal','pop')],afatalrate=10000*F1$afatal/F1$pop)

#find state which changes the law
F1_stat_t <- F1 %>% group_by(state,jail) %>% summarise(n=n())
jail_stat <-  unique(F1_stat_t[F1_stat_t$n!=7,]$state) #ct85 nv83 oh83-86 or84 sc83 ut83
jail_data <- F1[F1$state %in% jail_stat,]

F2 <- F1
F2$income <- log(F2$income)
F2$drinkage[round(F2$drinkage)!=F2$drinkage]<-c(19,18,19,20,19,20,20,18,20,20,19,19,20,19,20,19,19,18,18,19,18,19,19)
F2$drinkage <- as.factor(F2$drinkage)
ty2 <- sapply(F2, class)
r <- {}
a <- 1
for (i in 1:dim(F2)[2]){
  if (ty2[i]=='factor'){
    r[a] <- i
    a <- a+1
  }
}
F3 <- F2[,-c(r)]
cor(F3)
```

## Data Visualization

Our primary concern was the effect of mandatory jail time. In general, the number of states with mandatory jail time is lesser than the amount without mandatory jail time. More specifically, in 1982, there were 9 states with mandatory jail time, this number increased to 15 in 1985 but declined back to 14 in 1987. Figure 1 plotted the afatal rate under different jail policies across the years. Quite contradictory to what we thought, the average afatal rate among states without mandatory jail seemed to be lower than that of states with mandatory jail. Thus in the latter part, we could use statistical methods to investigate this correlation further.

```{r,fig.height=2,fig.cap = "afatal rate under different jail policies over 1982-1988", fig.align = "center",fig.pos = 'H'}
ggplot(F2,aes(x=year,y=afatalrate))+
  geom_boxplot(aes(fill=jail))+
  scale_fill_manual(values = c("#99CCFF","#FF9999"))
```
For majority states, they kept their law on mandatory jail the same from 1982-1988. We might also want to consider what might happen if the law changed. Out of 48 states, there were six states which changed their laws, and in Figure 2, each state was plotted individually against the year, with the enforcement year labeled. There was no clear pattern in whether having mandatory jail will lead to a decrease or increase in afatal rate. 

```{r,fig.pos = 'H',fig.height=3,fig.cap = "afatal rate of 6 states which changed their laws", fig.align = "center"}
#figure of 
theme_update(plot.title = element_text(hjust = 0.5,size = 8,face='bold'),axis.text=element_text(size = 6),axis.title=element_text(size=6,face="bold"))
#ct
cols <-rep("#E0E0E0",6)
cols[1]<-c("#99CCFF")
names(cols)=jail_stat
plot1<-ggplot(jail_data,aes(x= year,y = afatalrate,color=state))+
  labs(title='Connecticut')+
  geom_line(aes(group = state),size=0.5,show.legend = FALSE)+
  scale_colour_manual(values = cols)+
  geom_vline(xintercept=4)+
  geom_text(aes(x=4, label="\njail=yes", y=1), colour="#FF9999", angle=90,size=2)+
  geom_text(aes(x=4, label="jail=no\n", y=1), colour="#FF9999", angle=90,size=2)
#nv
cols <-rep("#E0E0E0",6)
cols[2]<-c("#99CCFF")
names(cols)=jail_stat
plot2<-ggplot(jail_data,aes(x= year,y = afatalrate,color=state))+
  labs(title='Nevada')+
  geom_line(aes(group = state),size=0.5,show.legend = FALSE)+
  scale_colour_manual(values = cols)+
  geom_vline(xintercept=2)+
  geom_text(aes(x=2, label="\njail=yes", y=1), colour="#FF9999", angle=90,size=2)+
  geom_text(aes(x=2, label="jail=no\n", y=1), colour="#FF9999", angle=90,size=2)
#oh
cols <-rep("#E0E0E0",6)
cols[3]<-c("#99CCFF")
names(cols)=jail_stat
plot3<-ggplot(jail_data,aes(x= year,y = afatalrate,color=state))+
  labs(title='Ohio')+
  geom_line(aes(group = state),size=0.5,show.legend = FALSE)+
  scale_colour_manual(values = cols)+
  geom_vline(xintercept=2)+
  geom_text(aes(x=2, label="\njail=yes", y=1), colour="#FF9999", angle=90,size=2)+
  geom_text(aes(x=2, label="jail=no\n", y=1), colour="#FF9999", angle=90,size=2)+
  geom_vline(xintercept=5)+
  geom_text(aes(x=5, label="jail=yes\n", y=1), colour="#FF9999", angle=90,size=2)+
  geom_text(aes(x=5, label="\njail=no", y=1), colour="#FF9999", angle=90,size=2)
#or
cols <-rep("#E0E0E0",6)
cols[4]<-c("#99CCFF")
names(cols)=jail_stat
plot4<-ggplot(jail_data,aes(x= year,y = afatalrate,color=state))+
  labs(title='Oregon')+
  geom_line(aes(group = state),size=0.5,show.legend = FALSE)+
  scale_colour_manual(values = cols)+
  geom_vline(xintercept=3)+
  geom_text(aes(x=3, label="\njail=yes", y=1), colour="#FF9999", angle=90,size=2)+
  geom_text(aes(x=3, label="jail=no\n", y=1), colour="#FF9999", angle=90,size=2)
#sc
cols <-rep("#E0E0E0",6)
cols[5]<-c("#99CCFF")
names(cols)=jail_stat
plot5<-ggplot(jail_data,aes(x= year,y = afatalrate,color=state))+
  labs(title='South Carolina')+
  geom_line(aes(group = state),size=0.5,show.legend = FALSE)+
  scale_colour_manual(values = cols)+
  geom_vline(xintercept=2)+
  geom_text(aes(x=2, label="\njail=yes", y=1), colour="#FF9999", angle=90,size=2)+
  geom_text(aes(x=2, label="jail=no\n", y=1), colour="#FF9999", angle=90,size=2)
#ut
cols <-rep("#E0E0E0",6)
cols[6]<-c("#99CCFF")
names(cols)=jail_stat
plot6<-ggplot(jail_data,aes(x= year,y = afatalrate,color=state))+
  labs(title='Utah')+
  geom_line(aes(group = state),size=0.5,show.legend = FALSE)+
  scale_colour_manual(values = cols)+
  geom_vline(xintercept=2)+
  geom_text(aes(x=2, label="\njail=yes", y=1), colour="#FF9999", angle=90,size=2)+
  geom_text(aes(x=2, label="jail=no\n", y=1), colour="#FF9999", angle=90,size=2)
grid.arrange(plot1, plot2,plot3, plot4, plot5, plot6,  nrow=2, ncol=3)
```
# Analysis
## Model building
In this experiment, it is a repeated measures design and observational study. Traffic fatalities data for 48 US states from 1982 to 1988 is available. A method that allows using data for more than two time periods and enables us to add control variables is the fixed effects regression approach. Fixed effects regression models eliminate omitted variable bias concerning potentially omitted variables that do not change over time, and they are often used to establish causal inference under specific circumstances. Thus, we employ fixed-effects models in this project to test the effect of mandatory jail sentences on afatal rate. Moreover, we will do model diagnostics and hypothesis testing.
```{r,results = "asis", message=FALSE, echo=FALSE, include=FALSE}
library(plm)
model <- plm(afatalrate ~ state+year+jail+drinkage+breath+service+spirits+unemp+income+beertax+
               baptist+mormon+dry+youngdrivers+milestot+gsp, 
                          index = c("state", "year"),
                          model = "within",
                          effect = "twoways",
                          data = F2)
lmodel <- lm(afatalrate ~ state+year+drinkage+breath+service+spirits+unemp+income+beertax+
               baptist+mormon+dry+youngdrivers+milestot+gsp+jail, data = F2)
summary(model)
length(model$coefficients)
model$coefficients
anova(lmodel)
```
### Fixed effects model:

$$Y_{{i,t}}=\mu +\alpha_{i}+\beta_{t}+\mathbf{\gamma} \mathbf{X}_{{it}}+\epsilon_{i,t}$$
\begin{align*}
&i \text{ denotes the index of states. al denotes Alabama, az denotes Arizona,..., wy denotes Wyoming. There are 48 states} \\
&t \text{ denotes the index of years. } t= 1982, 1983,\cdots, 1988.\\
&Y_{i,t} \text{ denotes the afatal rate in state i during yeat t}; \mu \text{ denotes the overall mean.}\\
&\alpha_{i} \text{ denotes the main effect of state i}; \beta_t \text{ denotes the main effect of year t.}\\
&\mathbf{X}_{{it}} \text{ is the time-variant } 1\times 16 \text{ regressor vector. }X_{it,1} \text{denotes whether the state i enforces mandatory jail in yeat t.}\\ 
&\text{If the state enforce mandatory jail in that year, } X_{it,1} = 1. \text{ Otherwise, } X_{it,1} = 0.\\
&\text{The explanation of }X_{it,k}, k = 2,\cdots,16 \text{ is shown in the Appendix 1.}\\
&\mathbf{\gamma} \text{ denotes the }{\displaystyle 16\times 1}\text{ matrix of parameters}; \varepsilon_{i,t} \text{ denotes random errors.}
\end{align*}

### Model assumptions
* \textit{\textbf{Independence}} error terms are independent with each other. In this experiment, we assume that each state at each time is independent with the same state at other years and is independent with other states at each year.
* \textit{\textbf{Normality}} Error terms are normally distributed.
* \textit{\textbf{Equal variance}} variances of error terms are all equal. $\sigma^2$ denotes variances of error terms.

Thus, error terms are independent and identically distributed random variables and are distributed as $Normal(0,\sigma^2).$ We will test normality assumptions and equal variance assumption in the Model Diagnostic section; in this project, we always assume that observations are independent and we will discuss this assumption in the discussion section.

### Fitted model

Since we are mainly interested in of mandatory jail sentence on the number of alcohol-involved vehicle fatalities, we only report the fitted value of $\gamma_1$. $\hat{\gamma}_1=0.226$. Other estimators are listed in Appendix 2.

## Model Diagnostics

```{r,fig.height = 2.5, fig.cap = "Model diagnostic plots. Left panel: Histogram of residuals. Middle panel: residual versus fitted values. Right panel: QQ plot with residuals",fig.align="center"}

par(mfrow=c(1,3))
hist(residuals(lmodel))
plot(lmodel,which = 1)
plot(lmodel,which = 2)

```
* \textit{\textbf{Normaility}} According to residual histogram, the distribution of residuals is left-skewed, which indicating that the model residuals are not normally distributed. The residuals QQ plot shows heavy-tail. Normality assumption is violated. 
* \textit{\textbf{Equal Variance}} The residual versus fitted values plot shows the sepraed of residuals are not even which indicates that equal variance assumption does not hold.  

## Casual inference

```{r, message=FALSE, echo=FALSE, include=FALSE}

F4 <- F2 %>%
      mutate(breath = ifelse(breath == "no",0,1),
            service = ifelse(service == "yes", 1, 0),
            jail = ifelse(jail == "yes", 1, 0))
F4$drinkage <- as.numeric(as.character(F2$drinkage))

F4 %>%
  group_by(jail) %>%
  summarise(n_obs = n(),
            mean_fatality = mean(afatalrate),
            std_error = sd(afatalrate) / sqrt(n_obs))


#calculate the mean for each covariate by the treatment status:


cov <- c('drinkage','breath','service','spirits','unemp','income','beertax',
               'baptist','mormon','dry','youngdrivers','milestot','gsp')

F4 %>%
  group_by(jail) %>%
  select(one_of(cov)) %>%
  summarise_all(funs(mean(., na.rm = T)))

lapply(cov, function(v) {
    t.test(F4[, v] ~ F4[, 'jail'])
})


# Propensity Score Estimation
m_ps <- glm(jail ~ drinkage+breath+service+spirits+unemp+income+beertax+
               baptist+mormon+dry+youngdrivers+milestot+gsp,
            family = binomial(), data = F4)


summary(m_ps)

# create propensity score data
prs_df <- data.frame(pr_score = predict(m_ps, type = "response"),
                     jail = m_ps$model$jail)

# draw the histogram of propensity score
labs <- paste("Mandatory jail:", c("yes", "no"))
p_s <- prs_df %>%
  mutate(jail = ifelse(jail == 1, labs[1], labs[2])) %>%
  ggplot(aes(x = pr_score)) +
  geom_histogram(color = "green") +
  facet_wrap(~jail) +
  xlab("Prob. of mandatory jail: before matching") +
  theme_bw()

# execute the match
F4_nomiss <- F4 %>%  # MatchIt does not allow missing values
  select(afatalrate, jail, one_of(cov)) %>%
  na.omit()

mod_match <- matchit(jail ~ drinkage+breath+service+spirits+unemp+income+beertax+
               baptist+mormon+dry+youngdrivers+milestot+gsp,
                     method = "nearest", data = F4)

dta_m <- match.data(mod_match)
dim(dta_m)


# create propensity score data after match
m_ps2 <- glm(jail ~ drinkage+breath+service+spirits+unemp+income+beertax+
               baptist+mormon+dry+youngdrivers+milestot+gsp,
            family = binomial(), data = dta_m)
prs_df2 <- data.frame(pr_score = predict(m_ps2, type = "response"),
                     jail = m_ps2$model$jail)
p_s2 <- prs_df2 %>%
  mutate(jail = ifelse(jail == 1, labs[1], labs[2])) %>%
  ggplot(aes(x = pr_score)) +
  geom_histogram(color = "green") +
  facet_wrap(~jail) +
  xlab("After matching") +
  theme_bw()


dta_m %>%
  group_by(jail) %>%
  select(one_of(cov)) %>%
  summarise_all(funs(mean))


lapply(cov, function(v) {
    t.test(dta_m[, v] ~ dta_m$jail)
})

#Estimating treatment effects
with(dta_m, t.test(afatalrate ~ jail))

c_t <- t.test(afatalrate ~ jail, data=dta_m)
p_value = round(unname(c_t[[3]]),digits=3)
mean1 = round(unname(c_t[[5]][1]),digits = 2)
mean2 = round(unname(c_t[[5]][2]),digits = 2)

#lm_ci <- lm(afatalrate~state+year+jail+drinkage+breath+service+spirits+unemp+income+beertax+
 #              baptist+mormon+dry+youngdrivers+milestot+gsp, data=dta_m)
#summary(lm_ci)
```

This project is an observational study, there is no randomization; for this kind of study, we used the propensity score method to make the causal inference

The Assumptions of this method:

* \textit{\textbf{Stable unit-treatment value}} The effect of mandatory jail sentence of one state at one year is on the alcohol-involved vehicle fatalities rate of that state at that year and does not spill over onto other units.

* \textit{\textbf{No latent variable}} This means that all confounding variables are observed.

* \textit{\textbf{Same range of propensities }} The range of propensities to be treated is the same for treated and control cases, even if the density functions have different shapes. 

These three assumptions are strong. However, to make causal inferences for observational data, we have to make some assumptions. With these assumptions, based on the theory of "The central role of the propensity score in observational
studies for causal inference" and following the processes of "R Tutorial 8: Propensity Score Matching", we employ propensity score and nearest-neighbor matching method to make a causal inference. Firstly, we estimate the propensity score through building a logit model where the outcome variable is a binary variable indicating mandatory jail law. Then, the propensity score for each observation could be calculated by the model. Through nearest neighbor matching method, we restrict the sample to observations within the region of common support, and then to divide the sample within the region of common support into five quintiles, based on the estimated propensity score, then 95% of the bias due to confounding will be eliminated. Finally, we use the t-test to test the effect of jail on afatal rate. The $p_{value}$ of the test is `r p_value`. Therefore, we are 95% confident that mandatory jail law causes a higher alcohol-involved vehicle fatalities rate.

```{r fig.height = 1.5, fig.cap = "Propensity Score Plot", fig.align = "center", fig.pos='H'}


require(gridExtra)
grid.arrange(p_s, p_s2, ncol=2)


```
 

## Trasformation and Hypothesis testing

### Transformation of response variable
```{r,results = "asis", message=FALSE, echo=FALSE, include=FALSE}
library(MASS)
boxcox(afatalrate ~ state+year+drinkage+breath+service+spirits+unemp+income+beertax+
               baptist+mormon+dry+youngdrivers+milestot+gsp+jail, data = F2)
F5 <- F2
F5$afatalrate <- log(F2$afatalrate)
hist(F2$afatalrate)
hist(log(F2$afatalrate))
hist((F2$afatalrate)^-1)
hist(sqrt(F2$afatalrate))
modelt <- lm(afatalrate ~ jail+state+year+drinkage+breath+service+spirits+unemp+income+beertax+
               baptist+mormon+dry+youngdrivers+milestot+gsp, data = F5)
summary(modelt)
co <- summary(modelt)$coefficients
co[2,1]+co[2,2]*qt(0.975,266)
co[2,1]-co[2,2]*qt(0.975,266)
```

In the fixed-effect model, some model assumptions do not hold. Thus, we consider transforming the response variables. Boxcox procedure, as well as the distributions, show that logarithm transformation seems to work the best in this case. Therefore, we chose logarithm transformation, and denoted the response variable as 'log(afatal)'. 
```{r,fig.height = 2, fig.cap = "Transformation of the response variable. Left panel: Boxcox procedure. Middle panel: The distribution of log(afatal). Right panel: The distribution of 1/afatal",fig.align="center",fig.pos='H'}
par(mfrow=c(1,3))
library(MASS)
boxcox(afatalrate ~ state+year+drinkage+breath+service+spirits+unemp+income+beertax+
               baptist+mormon+dry+youngdrivers+milestot+gsp+jail, data = F2)
hist(log(F2$afatalrate))
hist((F2$afatalrate)^-1)
```
### Model building and diagnostics

We will build a new fixed-effect model and the only difference between this model(model 2) and the previous model(model 1) is the response variable. New fixed-effect model:$log(Y_{{i,t}})=\mu +\alpha_{i}+\beta_{t}+\mathbf{\gamma} \mathbf{X}_{{it}}+\epsilon_{i,t}$. The notations of variables and model assumptions are same with model 1. In model 2, fitted value of $\gamma_1$ is $\hat{\gamma}_1=0.326$. Other estimators are listed in Appendix 3.  
In Figure 6, the residuals versus fitted values plot show the spread of residuals are even, which indicates that the equal variance assumption holds. The residuals QQ plot seemS to be reasonable, and the Normality assumption holds. The residuals versus Leverage plot shows that there is no influential outlier. Thus, the normality assumption and equal variance assumption hold. We assume that observations are independent, and we will discuss it later. Then, all assumptions hold.
```{r,fig.pos='H',fig.height = 2.5,fig.cap = "Model diagnostic plots. Left panel: residual versus fitted values. Middle panel: QQ plot with residuals. Right panel: Residuals versus Leverage plot",fig.align="center"}
par(mfrow=c(1,3))
plot(modelt,which = 1)
plot(modelt, which = 2)
plot(modelt,which = 5)
```
### Hypothesis testing 

#### The effect of mandatory jail sentence on the alcohol-involved vehicle fatalities rate
```{r,results = "asis", message=FALSE, echo=FALSE, include=FALSE}
summary(model)
length(model$coefficients)
model$coefficients
anova(lmodel)
summary(lmodel)
lmodel$coefficients
length(lmodel$coefficients)
48*7-length(lmodel$coefficients)
```
The primary question of the report is to test whether the mandatory jail sentence influences the alcohol-involved vehicle fatalities rate. We will test whether jail and log(afatal) are related first.
The null hypothesis is:
$$ H_0: \gamma_1 = 0, \text{ which means jail has no effect on log(afatal).}$$
against the alternative hypothesis
$H_a: \gamma_1 \neq 0, \text{ which means jail has a effect on log(afatal).}$ The $1-\alpha$ confidence interval for $\gamma_1$ is $\hat{\gamma}_1\pm t(1-\frac{\alpha}{2},266)*s(\hat{\gamma_1}).$ $t(1-\frac{\alpha}{2},266)$ is the $(1-\frac{\alpha}{2})$th percentile of t-distribution with 226 degrees of freedom. $s(\hat{\gamma_1})$ denotes standard errors of $\hat{\gamma_1}.$ In this model, 95% confidence interval for $\gamma_1$ is $[0.09,0.56].$ Since zero is not contained in the 95% confidence interval, we 95% confident that jail has association with log(afatal). Thus, it is likely that there is an association between mandatory jail law and afatal, which is consistent with the result of causual inference.

#### The main effect of states
Another question of interest is whether states have an association with afatal. Thus, we will test whether states and afatal is correlated. The 
$$H_{0}:\alpha_{al}=\alpha_{az}=...=\mu_{wy}=0; \text{ which means states have no effect on log(afatal).}$$
against the alternative hypothesis $H_{a}$:Not all $\alpha_{al},\alpha_{az},...,\mu_{wy}$ equal zero which means states have effect on log(afatal). The test statistics is F ratio:
$F^* = \frac{\frac{SSE(R)-SSE(F)}{df_R-df_F}}{\frac{SSE(F)}{df_F}}$
$SSE(F)$ denotes the error sum of squares(SSE) of model 2 and SSE(R) denotes SSE of model 3 that includes all variables contained in model 2 except for states; $df_R$ denotes the degrees of freedom of $SSE(R)$ and $df_F$ denotes the degrees of freedom of $SSE(F)$.  
At significant level $\alpha$, under $H_0$, $F^*\sim F(df_R-df_F,df_F)$. Thus, if $P(F(df_R-df_F,df_F)>F^*) < \alpha$, the null hypothesis is rejected at level $\alpha$. In the project, $P(F(df_R-df_F,df_F)>F^*)<0.05$ and $H_0$ is rejected at significant level 0.05. Therefore, it is likely that states and log(afatal) are related.

# Suggestions to Policymakers

In this experiment, we are surprised to find that having mandatory jail sentences will increase traffic fatalities. It is highly statistically significant. One of the potential reasons behind this finding is that the driver may act even more unlawfully to avoid the mandatory jail sentence and lead to the larger possibilities of traffic fatalities. This finding provides evidence for lifting the mandatory jail sentence to decrease traffic fatalities. 

# Discussion 
* \textit{\textbf{Independences }} In this project, we assume all observations are independent. However, the law or the rate of alcohol-involved vehicle fatalities of one state at one year may be highly correlated with the situation of that state at other years and with the situation of other states in the same year. These two points can explain that the data between the years and the states may not meet the assumption of independence. If we find a method to analyze data without an independence assumption, we will try to solve this problem.
* \textit{\textbf{Transformation of response variable }} When equal variance assumption and normality assumption are both violated, we transform the response variable. However, the meaning of the response variable is changed, and hypothesis testing could only test the relation between jail and log(afatal) instead of the relationship between jail and afatal. But we do not find better approaches to solve this problem. 

# Reference

1. Alcohol Policies and Highway Vehicle Fatalities, Christopher J. Ruhm, July 1995
2. Digest of State Alcohol-Highway Safety Related Legislation
3. U.S. history of alcohol minimum purchase age by state, https://en.wikipedia.org/wiki/U.S._history_of_alcohol_minimum_purchase_age_by_state
4. R Tutorial 8: Propensity Score Matching, https://sejdemyr.github.io/r-tutorials/statistics/tutorial8.html?nsukey=7wLCsKIO6LiBgcAs3XJCGQN3uY7dfeOQDebjUxkOpk%2Bnn3ieZM51tf3tajom5%2FTNSZk3gkwZl14ES6Sx7k627urS9I%2FXKBcWZNrCbFlsArlSctYCdqa%2BcATOvajp8KHKWsqHMnTo9wRiBKU%2FJ7pdPaPECcTMsWN52scsdfWlA%2FG77g0grEY0qYddfnFvp43Zz9o7599xvQyQmFweoGEBOA%3D%3D
5. Variable selection for propensity score models, Brookhart MA, et.al.,2007
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1513192/
6. Causal inference with observational data, Austin Nichols,2007
7. The central role of the propensity score in observational studies for causal effects

# Appendix

## Appendix 1: Explanation of predictor
\begin{itemize}
\item $X_{it,2},X_{it,3},X_{it,4}$ denote whether minimum legal drinking age in state i at yeat t was 19, 20, 21, respectively. There are dummy variables.
\item $X_{it,5},X_{it,6}$ denote whether preliminary breath test law or mandatory community service was enforced in state i at yeat t, respectively. There are dummy variables.
\item $X_{it,7},X_{it,8},\cdots,X_{it,16}$ respectively denote spirits consumption, unemployment rate, per capita personal income in 1987 dollars, tax on case of beer, percent of southern baptist, percent of mormon, percent residing in “dry” countries, percent of drivers aged 15–24, total vehicle miles (millions), GSP rate of change.
\end{itemize}

## Appendix 2: Fitted values
```{r}
summary(model)
```
***

Team ID: Course project group 13

Name (responsibilities): Zheng Gu (Hypothesis Testing, Model Diagnostics)

Name (responsibilities): Jieyun Wang (Background, Descriptive Analysis, Polish Report)

Name (responsibilities): Siyao Wang (Causal Inference, Model Fitting, Polish Report)

Name (responsibilities): Zhi Zhang (Causal Inference, Polish Report)

Github: https://github.com/jynwang/STA207Project.git
